Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2016-08-29 WSSSPE3/report/JORS-reviews-responses.txt

The paper lays out it's structure at the start, but each section could end
with additional signposts ("having done x, and before doing y, we will
review z"). I think those help the reader know where they are in a long
document.

RESPONSE:

Time has passed since the submission was prepared. I am conscious that my
late review played a role here. However it does create an opportunity to
check in with each of the groups and see if any progress has been made. For
example, it seems as if the effort to create the "best practices" paper has
flagged (at least nothing has been done since November 2015 at the gdoc
linked from the paper. Perhaps ask the conveners/contacts of each working
group if they have any updates?

RESPONSE:  We feel that this report should stick with where things were at
WSSSPE3, since we otherwise will never be done updating it. But for the
reviewer’s benefit, some groups have continued to work and have made good
progress, others have basically stopped work, and some are in the middle.
Some discussion of this can be found in
http://www.slideshare.net/danielskatz/summary-of-wssspe-focused-on-wssspe3

@@

General Comments:

The authors have provided a dense, lengthy but complete record of the
WSSSPE3 conference.  It follows on the WSSSPE2 report published in JORS, and
is balanced, trustworthy, and valuable. The authors’ limited
editorializing is simultaneously welcome in a historical report and renders
the lengthy document difficult to read or interpret since it lacks a strong
narrative form. To put that a bit more harshly, the manuscript feels like a
network graph with a goal of encompassing all the historical facts around
WSSSPE3. I think this stems from having a dual goal of tracing accurately
what happened at this conference and trying to encode action items and
practical outcomes into an authoritative document about what WSSSPE as an
entity is trying to become. I would recommend that this aspect of the
manuscript could be improved. Most of my concern is centered on Section 5.

Section 5 “Working Groups” beings without any narrative explaining where
these working groups came from and how they might have been culled from a
list of potential/actual working groups. The 10 working groups do not
correlate precisely with the “18 potential topics” on page 3 nor the 6
categories superimposed on the 18 topics. I suppose the working groups could
have been drawn from them though.  So this sets an unanswered question: were
these working groups decided upon by WSSSPE3 conference members or by a
WSSSPE body before the event? (Aside: does a WSSSPE entity exist today? Did
it exist before WSSSPE1? After WSSSPE2?)

RESPONSE: This is a very good observation.  We have added a few paragraphs to
the start of Section 5 to explain how the 10 initial working groups came about.
Regarding the second part of this comment, what has happened after WSSSPE3, we
feel that this report should stick with where things were at WSSSPE3, since
we otherwise will never be done updating it. But for the reviewer’s benefit,
some groups have continued to work and have made good progress, others have
basically stopped work, and some are in the middle.  Some discussion of this
can be found in http://www.slideshare.net/danielskatz/summary-of-wssspe-focused-on-wssspe3

@@

Before Section 5 the last time the reader was introduced to the working
group model was on page 3, with the rather diffuse text, “The first call
for participation requested lightning talks, where each author could make a
brief statement about work that either had been done or was needed, with the
goal of contributing to the discussion of one or more working groups.” I
think what is lacking is some threading through of where these working
groups came from and how they are the focus of WSSSPE# going forward.

RESPONSE: The new paragraphs at the start of Section 5 should resolve
this concern.

@@

It is also distracting that each subsection of Section 5 represents a
subset/working group speaking with that voice, e.g., Section 5.3, “At
WSSSPE3 our group was interested in better understanding successful pathways
for scientific software to “transition” from grant-funded research
projects to industry sponsorship.” I think it makes clear sense for
individual appendices to speak with a “group” voice, but that the main
document should speak with a single authoritative voice.

RESPONSE:

To fix my perceived problem with Section 5, some background would need to be
added and some of the summary material about working groups that is
presently in the Conclusions section could be moved to introduce the working
groups better.

RESPONSE:  Again, the new paragraphs at the start of Section 5 provide
the requested background.

### 2016-08-29 WSSSPE3/report/WSSSPE3_review.tex
more work in addressing reviewer comments

>>>>

After the keynote and lightning talks, the workshop facilitators led an exercise in which the
attendees first suggested additional topics beyond those in the call for participation (as listed in
\S\ref{sec:preworkshop}).  One topic that was introduced at this point was ``Building Sustainable
User Communities for Scientific Software.''  The full set of topics were then placed on
flipchart-sized pieces of paper around the walls, and the attendees voted on which topics
they were strongly interested in working on at WSSSPE3, and which topics they were 
generally interested in contributing to, but not as strongly.  This led to some topics being
taken out of the mix for the rest of the workshop, since not enough people wanted to contribute
to them to lead to a useful discussion.  Additionally, some topics were brought together by the
participants who felt they were closely linked and that a group could address multiple of them
in a single discussion.

After this, the attendees broke up into small working groups to discuss
the remaining topics during most of the remaining 1 1/2 days.  A high-level summary
of each topic and group's work can be found in the subsections of this section, and for
all but one group, 
more detailed notes on each group's discussions can be found in Appendices~\ref{sec:appendix_best_practices}--\ref{sec:appendix_user_community}.

Midway through the first afternoon and between the two days, each group
had a chance to talk about what progress they had made.  As discussed below (in \S\ref{sec:legacy}), the group that formed to discuss Legacy Software dissolved after the first session
as group members left to join other groups.
In addition, in the morning of the
second day, a small set of reviewers/advisors (external to the groups but chosen by the group
members) visited each group to listen to what they group was planning and to provide feedback.
